---
title: "Report VirusWarn-SC2 `r Sys.Date()`"
author: "VirusWarn"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
params:
  input: ""
  metadata: ""
  clusters: "vocal-alerts-clusters-summaries-all.csv"
  alert_samples: "vocal-alerts-samples-all.csv"
  moc: ""
  strict: "n"
  data: ""
  version: ""
---

<style>
  body .main-container {
      max-width: 1600px;
  }
</style>


#### **List of Abbrevations**

Sequences are sent to a category (Pink, Red, Orange, Grey) according to their mutation profile. 
VirusWarn counts the amount of mutations in 3 categories:

**MOC - Mutation of Concern**

Those are mutations that have been identified previously as changing the potential of the virus. 
The table of all MOC is at the bottom in the [MOC table](#item-moc).

**ROI - Region of Interest**

Those are important regions of the gene like for example Antigenic sites or Receptor-binding sites. 
The table of all ROI is at the bottom in the [ROI table](#item-roi).

**PM - Private Mutation**

Mutations that are neither a MOC nor a ROI.


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE)
```

```{r import, include=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
library(stringi)
library(lubridate)
library(tools)

library(seqinr)

library(ggplot2)
library(plotly)

library(DT)
```


```{r var, include=FALSE}
spike_len <- 1273

undefined <- c(
  'UNASSIGNED', 'Unassigned', 'unassigned', 
  'NONE', 'None', 'none', 
  'OTHER', 'Other', 'other'
)
```

```{r functions clusters, include=FALSE}
make_table_clusters <- function(clusters, color){
  clusters <- clusters %>%
    filter(alert_level == color) %>%
    select(
      n_samples, Lineages, s_moc_M.avg, s_moc_D.avg, s_roi_M.avg, s_roi_D.avg, 
      ListFrequentMutations_gt30perc, cluster_ID_in_alert_level
    ) %>%
    `colnames<-`(c("Samples", "Lineages", "MOC Sub Avg", "MOC Del Avg", "ROI Sub Avg", "ROI Del Avg", "Frequent Mutations", "Cluster ID"))
  
  result_table <- datatable(
    clusters, 
    class = 'cell-border stripe', 
    options = list(
      dom = 'lftp', 
      lengthMenu = c(5, 10, 20, 50),
      pageLength = 5,
      searchHighlight = TRUE,
      autoWidth= TRUE,
      columnDefs = list(list(width = 350, targets = c(7))),
      initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")
    )
  )  
  
  return(result_table)
}
```

```{r functions lollipop, include=FALSE}
make_lollipop <- function(table, color){
  lineage <- c()
  for(i in table$ID){
    new_lineage <- samples[samples$ID == i, "LINEAGE"]
    lineage <- c(lineage, new_lineage[[1]])
  }
  
  lineage_df <- table
  lineage_df$lineage <- lineage
  lineage_df$group <- "All Sequences"
  lineage_unique <- unique(lineage)
  lineage_len <- length(unique(lineage))
  
  df_list <- list(lineage_df)
  for (i in lineage_unique){
    new_df <- subset(lineage_df, lineage == i)
    new_df$group <- i
    df_list <- c(df_list, list(new_df))
  }
  
  groups <- c()
  data_list <- list()
  for (df in df_list){
    mut_freq <- rep(0, spike_len)
    mut_list <- vector(mode = "character", length = spike_len)
    mut_list_unique <- vector(mode = "character", length = spike_len)
    mut_list_unique_len <- vector(mode = "integer", length = spike_len)
    for (i in df$Mutations){
      mutations <- str_split_1(i, ",")
      for (j in mutations) {
        ind <- as.integer(str_sub(j, 2, -2))
        mut_freq[ind] <- mut_freq[ind] + 1
        mut_list[ind] <- paste(j, mut_list[ind], sep=" ")
      }
    }
    
    pos <- 1
    for (i in mut_list){
      split <- str_split_1(i, " ")
      p <- 1
      for (k in split){
        if (grepl("\\*", k) == TRUE){
          split[p] <- paste("\\", k, sep = "")
          print(split[k])
        }
        p <- p + 1
      }
      unique_mut <- stri_unique(split)
      for (j in unique_mut[-length(unique_mut)]){
        count <- str_count(i, j)
        mut_list_unique[pos] <- paste(mut_list_unique[pos], "<br>", toString(count), j)
        mut_list_unique_len[pos] <- mut_list_unique_len[pos] + 1
      }
      pos <- pos + 1
    }
    
    data_temp <- data.frame(x = seq(1, spike_len), y = mut_freq, list = mut_list_unique, list_len = mut_list_unique_len, group = df$group[1])
    groups <- c(groups, df$group[1])
    data_list <- c(data_list, list(data_temp))
  }
  
  all_false <- rep(FALSE, length(groups)*2)
  button_list <- list()
  for (i in 1:length(groups)) {
    visibility_state <- all_false
    visibility_state[(2 * i - 1):(i + i)] <- TRUE
    button_list[[i]] <- list(
      label = groups[i],
      method = "update",
      args = list(list(visible = visibility_state), list(title = groups[i]))
    )
  }
  
  p <- ggplot() +
    theme_light() +
    theme(
      panel.grid.major.x = element_blank(),
      panel.border = element_blank(),
      axis.ticks.x = element_blank()
    ) +
    xlab("Amino Acid Position") +
    ylab("Number of samples") +
    xlim(c(0,spike_len))
  
  for (i in 1:length(groups)) {
    d <- data_list[[i]][which(data_list[[i]]$y > 0),]
    row.names(d) <- seq(nrow(d))
    p <- p + 
      geom_segment(data = d, aes(x=x, xend=x, y=0, yend=y), color="darkgrey") +
      geom_point(
        data = d, 
        aes(
          x=x, y=y, 
          text=paste0(
          "<b>AA Position: </b>", x, 
          "<br><b># of samples: </b>", y, 
          "<br><b># of substitutions: </b>", list_len, list
          )
        ),
        color=color,
        size=3
      )
  }
  
  lollipop <- ggplotly(p, tooltip = c("text")) %>% layout(
    updatemenus=list(
      list(
        active=0, 
        x=0, 
        y=1.1,
        xanchor="left",
        yanchor="bottom",
        buttons=button_list
      )
    )
  )
  
  return(lollipop)
}
```

```{r functions heatmap, include=FALSE}
make_heatmap <- function(metadata, table){
  df_mutations_filtered <- merge(metadata, table, by="ID") %>%
    select(SAMPLING_DATE.x, Mutations) %>%
    unnest(Mutations) 
    
  df_week <- merge(metadata, table, by="ID") %>% 
    select(SAMPLING_DATE.x) %>%
    group_by(SAMPLING_DATE.x) %>%
    summarize(count = n())
  
  df_mutations_merged_all <- merge(df_mutations_filtered, df_week, by="SAMPLING_DATE.x") %>%
    select(Mutations, SAMPLING_DATE.x, count)
  
  df_mutations_merged <- df_mutations_merged_all %>%
    drop_na(SAMPLING_DATE.x)
  
  if (nrow(df_mutations_merged_all) - nrow(df_mutations_merged) != 0){
    df_na <- df_week %>% drop_na(SAMPLING_DATE.x)
    cat(paste0(sum(df_week$count) - sum(df_na$count), " Sequences were dropped from the heatmap due to the wrong format of the Sampling Date!"))
  }
  
  names(df_mutations_merged)[1] <- "aa_profile"
  names(df_mutations_merged)[3] <- "count_total"
  
  df_mutations_grouped <- suppressMessages(df_mutations_merged %>% 
    group_by(SAMPLING_DATE.x, count_total, aa_profile) %>% 
    summarize(freq = n()) %>%
    mutate(perc = round(freq/count_total*100, digits = 2)) %>% 
    ungroup())
  
  pos <- c()
  for (i in df_mutations_grouped["aa_profile"]){
    # get position of mutation
    ind <- as.integer(str_sub(i, 2, -2))
    pos <- c(pos, ind)
  }
  df_mutations_grouped$position <- pos
  
  df_mutations_sorted <- df_mutations_grouped %>%
    arrange(-position)
  
  plot.data <- df_mutations_sorted %>% 
    select(SAMPLING_DATE.x, aa_profile, freq, perc)
  
  year_weeks <- as.Date(paste0(plot.data$SAMPLING_DATE.x, "-1"), format="%Y-%U-%u")
  week_sequence <- seq(min(year_weeks), max(year_weeks), by = "week")
  full_seq <- format(week_sequence, "%Y-%U")
  full_seq <- full_seq[full_seq %in% unique(plot.data$SAMPLING_DATE.x)]
  missing <- setdiff(full_seq, plot.data$SAMPLING_DATE.x)
  missing <- missing[!grepl("-00$", missing)]
  
  if (length(missing) != 0){
    for (i in 1:length(missing)){
      new_row <- list(SAMPLING_DATE.x = missing[[i]], aa_profile = NA, freq = NA, perc = NA)
      plot.data <- rbind(plot.data, new_row)
    }
  }
  
  aa_tick <- unique(plot.data$aa_profile)
  date_tick <- unique(plot.data$SAMPLING_DATE.x)
  
  heatmap <- plot_ly(
    plot.data, x=~SAMPLING_DATE.x, y=~aa_profile, z=~perc, 
    reversescale=TRUE, 
    xgap=1.5, ygap=1.5, 
    type="heatmap", 
    colorbar=list(
      title=('<b>Proportion</b>')), 
      hoverinfo='text', 
      text=~paste0("<b>Week: </b>", .data$SAMPLING_DATE.x, "<br><b>Substitution: </b>", .data$aa_profile, "<br><b>% of samples: </b>", .data$perc)
    ) %>% 
    layout(
      plot_bgcolor="#F8F8F8", 
      xaxis=list(nticks=length(date_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(nticks=length(aa_tick), title="<b>Substitutions</b>", showgrid=FALSE, categoryorder="trace"),
      legend=list(orientation='v', xanchor="right", yanchor="top")
    ) 
  
  df_mutations_bar <- merge(metadata, table, by="ID") %>%
    select(SAMPLING_DATE.x, Mutations)
  
  bar.data <- df_mutations_bar %>%
    select(SAMPLING_DATE.x) %>%
    group_by(SAMPLING_DATE.x) %>%
    summarize(count = n())
  
  if (length(missing) != 0){
    for (i in 1:length(missing)){
      new_row <- list(SAMPLING_DATE.x = missing[[i]], count = 0)
      bar.data <- rbind(bar.data, new_row)
    }
  }
  
  bar.data <- bar.data %>% 
    arrange(SAMPLING_DATE.x)
  
  barplot <- plot_ly(
    bar.data, x=~SAMPLING_DATE.x, y=~count, 
    type = "bar",
    marker = list(color = "grey"),
    hoverinfo='text',
    hovertext=~paste0("<b>Week: </b>", .data$SAMPLING_DATE.x, "<br><b># of samples: </b>", .data$count)
  ) %>%
    layout(
      xaxis=list(nticks=length(date_tick), title='<b>Week (Collection Date)</b>', showgrid=FALSE),
      yaxis=list(title="<b>Number of <br> sequences</b>", showgrid=FALSE)
    )
  
  heatmap <- subplot(
    heatmap, barplot, 
    nrows = 2, heights = c(0.8, 0.2), 
    shareX = TRUE, titleY = TRUE
  )
  
  return(heatmap)
}
```

```{r functions table, include=FALSE}
prepare_table <- function(table){
  for (i in table["ID"]) {
    new <- str_remove_all(i, "%space%")
    table["ID"][table["ID"] == i] <- new
  }
  
  names(table)[7] <- "Mutations"
  names(table)[8] <- "Deletions"

  table_copy <- table
  
  for (i in table["Mutations"]){
    new <- str_split(i, ",")
    table["Mutations"][table["Mutations"] == i] <- new
  }
  return(list(table = table, table_copy = table_copy))
}
  
make_table_samples <- function(table_copy){
  for (i in table_copy["Mutations"]){
    new <- gsub(",([[:alpha:]])", ", \\1", i)
    table_copy["Mutations"][table_copy["Mutations"] == i] <- new
  }
  
  names(table_copy)[7] <- "Substitutions"

  names(table_copy)[4] <- "MOC"
  names(table_copy)[5] <- "ROI"
  names(table_copy)[6] <- "PM"
  
  col_order <- c("ID", "LINEAGE", "MOC", "ROI", "PM", "Substitutions", "Deletions")
  table_copy <- table_copy[, col_order]
  
  result_table <- datatable(
    table_copy, 
    class = 'cell-border stripe', 
    options = list(
      dom = 'lftp', 
      lengthMenu = c(5, 10, 20, 50),
      pageLength = 5,
      searchHighlight = TRUE,
      autoWidth= TRUE,
      columnDefs = list(list(width = 350, targets = c(6,7))),
      initComplete = JS(
      "function(settings, json) {",
      "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
      "}")
    )
  )  
  
  return(result_table)
}
```

```{r data, include=FALSE}
if (file_ext(params$input) == 'fasta'){
  input <- read.fasta(params$input)
} else {
  input <- read.delim(params$input, sep = ",")
}

clusters <- read.delim(params$clusters, sep=',')
samples <- read.delim(params$alert_samples, sep = ',')

moc <- read.delim(params$moc)

if(params$metadata != ''){
  metadata <- read.delim(params$metadata)
}
```

```{r data groups, include=FALSE}
pink_table <- filter(samples, alert_level == 'pink') %>%
  mutate(
    s_moc = s_moc_M + s_moc_D + s_moc_I, 
    s_roi = s_roi_M + s_roi_D + s_roi_I, 
    s_pm = s_pm_M + s_pm_D + s_pm_I
  ) %>%
  select(
    ID, SAMPLING_DATE, LINEAGE, s_moc, s_roi, s_pm, 
    ListMutationsSelected_M, ListMutationsSelected_D,
    nMutationsTotal_M, nMutationsTotal_D, nMutationsTotal_I 
  )

red_table <- filter(samples, alert_level == 'red') %>%
  mutate(
    s_moc = s_moc_M + s_moc_D + s_moc_I, 
    s_roi = s_roi_M + s_roi_D + s_roi_I, 
    s_pm = s_pm_M + s_pm_D + s_pm_I
  ) %>%
  select(
    ID, SAMPLING_DATE, LINEAGE, s_moc, s_roi, s_pm, 
    ListMutationsSelected_M, ListMutationsSelected_D,
    nMutationsTotal_M, nMutationsTotal_D, nMutationsTotal_I
  )

orange_table <- filter(samples, alert_level == 'orange') %>%
  mutate(
    s_moc = s_moc_M + s_moc_D + s_moc_I, 
    s_roi = s_roi_M + s_roi_D + s_roi_I, 
    s_pm = s_pm_M + s_pm_D + s_pm_I
  ) %>%
  select(
    ID, SAMPLING_DATE, LINEAGE, s_moc, s_roi, s_pm, 
    ListMutationsSelected_M, ListMutationsSelected_D,
    nMutationsTotal_M, nMutationsTotal_D, nMutationsTotal_I
  )

grey_table <- filter(samples, alert_level == 'grey') %>%
  mutate(
    s_moc = s_moc_M + s_moc_D + s_moc_I, 
    s_roi = s_roi_M + s_roi_D + s_roi_I, 
    s_pm = s_pm_M + s_pm_D + s_pm_I
  ) %>%
  select(
    ID, SAMPLING_DATE, LINEAGE, s_moc, s_roi, s_pm, 
    ListMutationsSelected_M, ListMutationsSelected_D,
    nMutationsTotal_M, nMutationsTotal_D, nMutationsTotal_I
  )
```



## Overview

The report was generated for the following input file:

```{r input, echo=FALSE, message=FALSE, warning=FALSE}
cat(paste0('> ', params$input))
```

```{r seq total, echo=FALSE, message=FALSE, warning=FALSE}
if (file_ext(params$input) == 'fasta'){
  cat(paste0('> Sequences total in fasta file: ', length(input)))
  cat(paste0('> Sequences aligned: ', nrow(samples)))
  cat(paste0('> Sequences skipped: ', length(input) - nrow(samples)))
} else {
  cat(paste0('> Sequences total in input file: ', nrow(input)))
}
```

```{r metadata, echo=FALSE, message=FALSE, warning=FALSE}
if (params$metadata != ""){
  cat(paste0('> Metadata was given for the following amount of sequences: ', nrow(metadata)))
} else {
  cat("No Metadata file was given.")
}
```

### Sequences per Group

```{r bar plot,  out.width="1000px", echo=FALSE, message=FALSE, warning=FALSE}
alert_colors <- c(grey = "lightgrey", orange = "orange", red = "red", pink = "deeppink")

if (params$metadata == ""){
  df_bar <- samples %>%
    group_by(alert_level) %>%
    summarize(count = n())
  
  if (nrow(df_bar) != 5){
    all_alert_levels <- c('grey', 'orange', 'red', 'pink')
    missing_alert_levels <- setdiff(all_alert_levels, df_bar$alert_level)
    for (level in missing_alert_levels){
      new_row <- data.frame(alert_level = level, count = 0)
      df_bar <- rbind(df_bar, new_row)
    }
  }
  
  order_levels <- c('grey', 'orange', 'red', 'pink')
  df_bar$alert_level <- factor(df_bar$alert_level, levels = order_levels)
  count_sum <- sum(df_bar$count)
  
  plot_ly(
    df_bar, x = ~0, y = ~count,
    type = 'bar', 
    color = ~alert_level,
    colors = alert_colors,
    hoverinfo = 'text',
    hovertext = ~paste0("<b>Group: </b>", alert_level, "<br><b># of sequences: </b>", count, "<br><b>Percentage: </b>", round(count/count_sum*100, digits = 2), "%")
  ) %>% 
    layout(
      xaxis = list(title = 'Alert Level'),
      yaxis = list(title = 'Count'), 
      barmode = 'stack'
    )
} else {
  names_all <- vector(mode = "character", length = length(samples$ID))
  df_names_all <- cbind(samples, names_all)
  
  df_week_all <- merge(df_names_all, metadata, by="ID", all.x = TRUE) %>% 
    select(alert_level, ID, SAMPLING_DATE.x)
    
  df_bar <- df_week_all %>%
    group_by(alert_level, SAMPLING_DATE.x) %>%
    summarize(count = n())
  
  order_levels <- c('grey', 'orange', 'red', 'pink')
  df_bar$alert_level <- factor(df_bar$alert_level, levels = order_levels)
  
  plot_ly(
    df_bar, x = ~SAMPLING_DATE.x, y = ~count, 
    type = 'bar', 
    color = ~alert_level,
    colors = alert_colors,
    hoverinfo='text',
    hovertext=~paste0("<b>Collection Date: </b>", SAMPLING_DATE.x, "<br><b>Group: </b>", alert_level, "<br><b># of sequences: </b>", count)
  ) %>% 
    layout(
      xaxis = list(title = 'Collection Date'), 
      yaxis = list(title = 'Count'), 
      barmode = 'stack'
    )
}
```




## Group <span style="color: deeppink;">PINK</span> {.tabset .tabset-fade}
#### *Sequences that are a VOC / VOI / VUM with 1 or more MOC substitutions or 1 or more MOC deletions*

```{r pink data, echo=FALSE, message=FALSE, warning=FALSE}
result <- prepare_table(pink_table)
pink_table <- result$table
pink_table_copy <- result$table_copy

cat(paste0('> Sequences in Group PINK: ', nrow(pink_table)))

if (any(pink_table$LINEAGE %in% undefined)){
  cat(paste0("\nINFO: There are ", sum(pink_table$LINEAGE %in% undefined), " sequences with no assigned lineage in the category PINK. For more information, please check the tab Samples or the drop down menu of the Lollipop plot for Unassinged/NONE/Other."))
}
```

### Clusters

```{r pink clusters, echo=FALSE, message=FALSE, warning=FALSE}
if ('pink' %in% clusters$alert_level){
  pink_clusters <- make_table_clusters(clusters, 'pink')
  pink_clusters
} else {
  cat("No clusters in Group PINK")
}
```


### Lollipop Plot

```{r pink lollipop, out.width="1000px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(pink_table) == 0){
  cat("No sequences in Group PINK")
} else {
  pink_lollipop <- make_lollipop(pink_table_copy, "deeppink")
  pink_lollipop
}
```

### Heatmap

```{r pink heatmap, out.width="1000px", out.height="800px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(pink_table) == 0){
  cat("No sequences in Group PINK")
} else if (params$metadata == ""){
  cat("The heatmap shows the frequency of mutations over time.")
  cat("No metadata file with collection dates was given, so no heatmap was generated.")
} else if (any(pink_table$ID %in% metadata$ID) == FALSE){
  cat("No metadata given for the sequences of Group PINK.")
} else {
  pink_heatmap <- make_heatmap(metadata, pink_table)
  pink_heatmap
}
```

### Samples

```{r pink table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(pink_table) == 0){
  cat("No sequences in Group PINK")
} else {
  pink_result_table <- make_table_samples(pink_table_copy)
  pink_result_table
}
```


## Group <span style="color: red;">RED</span> {.tabset .tabset-fade}
#### *Sequences with 3 or more MOC substitutions and 2 or more MOC deletions and at least 1 PM substitution that are not PINK*
#### *Sequences with 2 or more MOC substitutions and 1 or more MOC deletions and 4 or more PM substitutions that are not PINK*

```{r red data, echo=FALSE, message=FALSE, warning=FALSE}
result <- prepare_table(red_table)
red_table <- result$table
red_table_copy <- result$table_copy

cat(paste0('> Sequences in Group RED: ', nrow(red_table)))

if (any(red_table$LINEAGE %in% undefined)){
  cat(paste0("\nINFO: There are ", sum(red_table$LINEAGE %in% undefined), " sequences with no assigned lineage in the category RED. For more information, please check the tab Samples or the drop down menu of the Lollipop plot for Unassinged/NONE/Other."))
}
```

### Clusters

```{r red clusters, echo=FALSE, message=FALSE, warning=FALSE}
if ('red' %in% clusters$alert_level){
  red_clusters <- make_table_clusters(clusters, 'red')
  red_clusters
} else {
  cat("No clusters in Group RED")
}
```

### Lollipop Plot

```{r red lollipop, out.width="1000px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(red_table) == 0){
  cat("No sequences in Group RED")
} else {
  red_lollipop <- make_lollipop(red_table_copy, "red")
  red_lollipop
}
```

### Heatmap

```{r red heatmap, out.width="1000px", out.height="800px", echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(red_table) == 0){
  cat("No sequences in Group RED")
} else if (params$metadata == ""){
  cat("The heatmap shows the frequency of mutations over time.")
  cat("No metadata file with sampling dates was given, so no heatmap was generated.")
} else if (any(red_table$ID %in% metadata$ID) == FALSE){
  cat("No metadata given for the sequences of Group RED.")
} else {
  red_heatmap <- make_heatmap(metadata, red_table)
  red_heatmap
}
```

### Samples

```{r red table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(red_table) == 0){
  cat("No sequences in Group RED")
} else {
  red_result_table <- make_table_samples(red_table_copy)
  red_result_table
}
```


## Group <span style="color: orange;">ORANGE</span> {.tabset .tabset-fade}
#### *Sequences with 2 or more MOC substitutions and 1 or more MOC deletions and 2 or more PM substitutions*

```{r orange data, echo=FALSE, message=FALSE, warning=FALSE}
if (params$strict == 'y'){
  cat("VirusWarn-SC2 runs in strict mode, the ORANGE level is not available.")
} else {
  result <- prepare_table(orange_table)
  orange_table <- result$table
  orange_table_copy <- result$table_copy
  
  cat(paste0('> Sequences in Group ORANGE: ', nrow(orange_table)))
  
  if (any(orange_table$LINEAGE %in% undefined)){
    cat(paste0("\nINFO: There are ", sum(orange_table$LINEAGE %in% undefined), " sequences with no assigned lineage in the category ORANGE. For more information, please check the tab Samples or the drop down menu of the Lollipop plot for Unassinged/NONE/Other."))
  }
}
```

### Clusters

```{r orange clusters, echo=FALSE, message=FALSE, warning=FALSE}
if (params$strict == 'y'){
  cat("VirusWarn-SC2 runs in strict mode, the ORANGE level is not available.")
} else if ('orange' %in% clusters$alert_level){
  orange_clusters <- make_table_clusters(clusters, 'orange')
  orange_clusters
} else {
  cat("No clusters in Group ORANGE")
}
```

### Lollipop Plot

```{r orange lollipop, out.width="1000px", echo=FALSE, message=FALSE, warning=FALSE}
if (params$strict == 'y'){
  cat("VirusWarn-SC2 runs in strict mode, the ORANGE level is not available.")
} else if (nrow(orange_table) == 0){
  cat("No sequences in Group ORANGE")
} else {
  orange_lollipop <- make_lollipop(orange_table_copy, "orange")
  orange_lollipop
}
```

### Heatmap

```{r orange heatmap, out.width="1000px", echo=FALSE, message=FALSE, warning=FALSE}
if (params$strict == 'y'){
  cat("VirusWarn-SC2 runs in strict mode, the ORANGE level is not available.")
} else if (nrow(orange_table) == 0){
  cat("No sequences in Group ORANGE")
} else if (params$metadata == ""){
  cat("The heatmap shows the frequency of mutations over the weeks.")
  cat("No metadata file with collection dates was given, so no heatmap was generated.")
} else if (any(orange_table$ID %in% metadata$ID) == FALSE){
  cat("No metadata given for the sequences of Group ORANGE.")
} else {
  orange_heatmap <- make_heatmap(metadata, orange_table)
  orange_heatmap
}
```

### Samples

```{r orange table, echo=FALSE, message=FALSE, warning=FALSE}
if (params$strict == 'y'){
  cat("FluWarnSystem runs in strict mode, the ORANGE level is not available.")
} else if (nrow(orange_table) == 0){
  cat("No sequences in Group ORANGE")
} else {
  orange_result_table <- make_table_samples(orange_table_copy)
  orange_result_table
}
```


## Group <span style="color: lightgrey;">GREY</span>

```{r grey data, echo=FALSE, message=FALSE, warning=FALSE}
result <- prepare_table(grey_table)
grey_table <- result$table
grey_table_copy <- result$table_copy

cat(paste0('> Sequences in Group GREY: ', nrow(grey_table)))

if (any(grey_table$LINEAGE %in% undefined)){
  cat(paste0("\nINFO: There are ", sum(grey_table$LINEAGE %in% undefined), " sequences with no assigned lineage in the category GREY. For more information, please check the tab Samples or the drop down menu of the Lollipop plot for Unassinged/NONE/Other."))
}
```

```{r grey table, echo=FALSE, message=FALSE, warning=FALSE}
if (nrow(grey_table) == 0){
  cat("No sequences in Group GREY")
} else {
  grey_result_table <- make_table_samples(grey_table_copy)
  grey_result_table
}
```



## Data Summary

The following file versions were used to generate the report:

Antibody Escape Score `escape_data_bloom_lab.csv`: 2021-10-27

Positive Selection Score `sergei_extracted.2021-05-17.csv`: 2021-05-17

```{r db version, echo=FALSE, message=FALSE, warning=FALSE}
if (params$data == 2021){
  cat("Lineage and sublineage relationship `lineage.all.tsv`: 2022-06-29\n")
  cat("ECDC assigned variants `ECDC_assigned_variants.csv`: 2022-06-29\n")
  cat("SARS-CoV-2 mutation information `table_cov2_mutations_annotation.tsv`: 2022-06-29\n")
} else if (params$data == 2022){
  cat("Lineage and sublineage relationship `lineage.all.tsv`: 2024-09-30\n")
  cat("ECDC assigned variants `ECDC_assigned_variants.csv`: 2024-09-30\n")
  cat("SARS-CoV-2 mutation information `table_cov2_mutations_annotation.tsv`: 2024-09-30\n")
} else {
  cat("Lineage and sublineage relationship `lineage.all.tsv`: Personalized file\n")
  cat("WHO assigned variants `WHO_assigned_variants.csv`: Personalized file\n")
  cat("SARS-CoV-2 mutation information `table_cov2_mutations_annotation.tsv`: Personalized file\n")
}
```


### Reference sequence

All results of VirusWarn-SC2 reports are calculated with the reference sequence **Wuhan-Hu-1 NC_045512.2**. 


<a id="item-moc"></a>

### MOC table

The following table shows the list of the Mutations of Concern that were used to generate the report.

```{r moc, echo=FALSE, message=FALSE, warning=FALSE}
moc_table <- moc %>% filter(gene == "S", type == "MutationOfConcern") %>%
  select(-nucleotide, -type) %>%
  mutate(gene = "SPIKE") %>%
  mutate(comments = paste(ID, comment, sep = ", ")) %>%
  select(-ID, -comment)

datatable(
  moc_table, 
  class = 'cell-border stripe', 
  options = list(
    dom = 'lftp', 
    lengthMenu = c(5, 10, 20, 50), 
    pageLength = 10,
    searchHighlight = TRUE,
    initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}")
  )
)
```


<a id="item-roi"></a>

### ROI table

The following table shows the list of Regions of Interest that were used to generate the report.

```{r roi, echo=FALSE, message=FALSE, warning=FALSE}
roi_table <- data.frame(
  gene = "SPIKE",
  nt_start = c(21602,21980,22295,22760),
  nt_end = c(21622,22036,22354,23065),
  aa_start = c(14,140,245,400),
  aa_end = c(220,158,264,501),
  func = "N-terminal domain, Antibody epitope"
)

datatable(
  roi_table, 
  class = 'cell-border stripe', 
  options = list(
    dom = 'lftp', 
    lengthMenu = c(5, 10, 20, 50),
    pageLength = 10,
    searchHighlight = TRUE,
    initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});",
    "}")
  )
)
```

**This report was generated with VirusWarn-SC2 `r params$version`.**
